---
title: "Importance Sampling"
author: "Qihun Yang"
date: "2023-04-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Ex 1
( 1 )
true value of the integral:
$$I=\frac{\sqrt{2}}{2}(e^{-\frac{9}{4}}+e^{-9})\approx0.07461577$$
```{r 3.4.0}
tr = (exp(-9/4)+exp(-9))*sqrt(2)/2;tr
```


( 3 )
```{r 3.4.1}
set.seed(1)
n = 1000
h4 <- function(x){
  exp(-0.5*(x-3)^2)+exp(-0.5*(x-6)^2)
}
```

This problem can be solved via general importance sampling or importance sampling concerning composition sampling technique:

-General Importance Sampling
```{r 3.4.2}
curve(h4,0,10)
```
From the curve of $h(x)$, I use the pdf of $N(4.5,4)$, which looks like this:
```{r 3.4.3}
norm4.5 <- function(x){
  return((1/(2*sqrt(2*pi)))*exp(-0.125*(x-4.5)^2))
}
curve(h4,0,10)
curve(norm4.5,add = TRUE)
```
This curve by and large fits $h(x)$.
```{r 3.4.4}
set.seed(1)
x4 <- rnorm(n,4.5,2)
X4 <- (dnorm(x4)*h4(x4))/norm4.5(x4)
I6 <- mean(X4);I6
e1 = 1.96*sqrt(mean((X4-tr)^2))/sqrt(n);e1
```

-Importance Sampling concerning composition sampling technique
$h(x)$ is very much like the mixture of pdfs of two distribution, $N(3,1)$ and $N(6,1)$. So I use:
$$g(x) \propto h(x)$$
The integral of $h(x)$ is:
\begin{aligned}
I&=\int_{-\infty}^{\infty}\exp \left(-\frac{1}{2}(x-3)^{2}\right)+\exp \left(-\frac{1}{2}(x-6)^{2}\right) dx \\
&=2\sqrt{2\pi} \\
&\approx5.013257
\end{aligned}

So $$g(x)=\frac{0.5}{\sqrt{2\pi}}\exp \left(-\frac{1}{2}(x-3)^{2}\right)+\frac{0.5}{\sqrt{2\pi}}\exp \left(-\frac{1}{2}(x-6)^{2}\right)$$
```{r 3.4.5}
g4 <- function(x){
  return(0.5/(sqrt(2*pi))*exp(-0.5*(x-3)^2)+0.5/(sqrt(2*pi))*exp(-0.5*(x-6)^2))
}
```
Sample $g(x)$ using the composition sampling technique:
```{r 3.4.6}
set.seed(1)
x4 <- c()
for (i in 1:n) {
  u <- runif(1,0,1)
  if(u<0.5){
    x4[i] <- rnorm(1,3,1)
  }else{
    x4[i] <- rnorm(1,6,1)
  }
}
X4 <- (dnorm(x4)*h4(x4))/g4(x4)
I6 <- mean(X4);I6
#Error bounds at 95% confidence level
e2 = 1.96*sqrt(mean((X4-tr)^2))/sqrt(n);e2
```
From error bounds, the second method is considered better.

## Ex 2
( 1 )

Suppose that generating $N$ standard normal random numbers can guarantee at least 1 number greater than 4.5 on average, that is:
$$E(n)=\theta N\geq1$$
$$N\geq294290.75$$
meaning that $N$ should be at least $294291$ to guarantee.

( 2 )

Estimating $\theta=P(X>4.5)=3.398 \times 10^{-6}$ is equivalent to calculate the integral:
$$I=\frac{1}{\sqrt{2\pi}}\int_{4.5}^{+\infty} e^{-\frac{1}{2}x^2}dx$$
```{r 3.5.1}
set.seed(1)
n = 1000
x5 <- -log(runif(n,0,1))+4.5
X5 <- dnorm(x5)/exp(-(x5-4.5))
theta <- mean(X5);theta
#Error bounds at 95% confidence level
e = 1.96*sqrt(mean((X5-3.398*10^(-6))^2))/sqrt(n);e
```
Note that the pdf of $W$ is $p_W(x)=e^{-(x-5)}$.
